Temporal Graph Cube

### 2023-01-05 TemporalGraphCube/old-version/TemporalGraphCube_full.tex
latex 1-5

\documentclass[10pt,journal,compsoc]{IEEEtran}
%\usepackage{ntheorem}
%\usepackage{algorithmic}
%\usepackage{cite}
%\usepackage{times}
\usepackage{amsmath,amsfonts,amssymb, amsthm, xspace}
\usepackage[linesnumbered,ruled,vlined]{algorithm2e}
\usepackage{graphicx, subfigure,stfloats}
\usepackage{epstopdf}
\usepackage{textcomp}
\usepackage{ragged2e, xcolor}
\usepackage{balance}
\usepackage{verbatim}
\usepackage{enumitem}
\usepackage{makecell}
\usepackage{url}
\usepackage{booktabs}

\long\def\comment#1{}
%\newcommand{\eop}{\hspace*{\fill}\mbox{$\Box$}}

%
% If IEEEtran.cls has not been installed into the LaTeX system files,
% manually specify the path to it like:
% \documentclass[10pt,journal,compsoc]{../sty/IEEEtran}

@@

\renewcommand{\baselinestretch}{0.97}

@@

	\author{Guoren~Wang, Yue Zeng, Rong-Hua Li, Hongchao Qin,\\ Xuanhua Shi, Yubin Xia, Xuequn Shang, Liang Hong
		\IEEEcompsocitemizethanks{
			\IEEEcompsocthanksitem G.~Wang, Y.~Zeng, R.-H.~Li, H.~Qin are with the Beijing Institute of Technology, China; Xuanhua Shi is with Huazhong University of Science and Technology, China; Yubin Xia is with Shanghai Jiao Tong University; Xuequn Shang is with Northwestern Polytechnical University; Liang Hong is with Wuhan University.

E-mail: wanggrbit@126.com, bruceez@163.com, lironghuabit@126.com, qhc.neu@gmail.com, xhshi@hust.edu.cn, xiayubin@sjtu.edu.cn,  shang@nwpu.edu.cn, hong@whu.edu.cn.
%			\IEEEcompsocthanksitem R.~Mao is with the Shenzhen University, Shenzhen 518060, China. E-mail: mao@szu.edu.cn.
		}
	}

@@

% The publisher's ID mark at the bottom of the page is less important with
% Computer Society journal papers as those publications place the marks
% outside of the main text columns and, therefore, unlike regular IEEE
% journals, the available text space is not reduced by their presence.
% If you want to put a publisher's ID mark on the page you can do it like
% this:
%\IEEEpubid{0000--0000/00\$00.00~\copyright~2015 IEEE}
% or like this to get the Computer Society new two part style.
%\IEEEpubid{\makebox[\columnwidth]{\hfill 0000--0000/00/\$00.00~\copyright~2015 IEEE}%
	%\hspace{\columnsep}\makebox[\columnwidth]{Published by the IEEE Computer Society\hfill}}
% Remember, if you use this you must call \IEEEpubidadjcol in the second
% column for its text to clear the IEEEpubid mark (Computer Society jorunal
% papers do not need this extra clearance.)

@@

% for Computer Society papers, we must declare the abstract and index terms
% PRIOR to the title within the \IEEEtitleabstractindextext IEEEtran
% command as these need to go into the title area created by \maketitle.
% As a general rule, do not put math, special symbols or citations
% in the abstract or keywords.
\IEEEtitleabstractindextext{%
	\begin{abstract}
		\justifying Data warehouse and OLAP (Online Analytical Processing) are effective tools for decision support on traditional relational data and static multidimensional network data. However, many real-world multidimensional networks are often modeled as temporal multidimensional networks, where the edges in the network are associated with temporal information. Such temporal multidimensional networks typically cannot be handled by traditional data warehouse and OLAP techniques. To fill this gap, we propose a novel data warehouse model, named \kw{Temporal\enspace Graph\enspace Cube}, to support OLAP queries on temporal multidimensional networks. Through supporting OLAP queries in any time range, users can obtain summarized information of the network in the time range of interest, which cannot be derived by using traditional static graph OLAP techniques. We propose a segment-tree based indexing technique to speed up the OLAP queries, and also develop an index-updating technique to maintain the index when the temporal multidimensional network evolves over time. In addition, we also propose a novel concept called \kw{similarity\enspace of\enspace snapshots} which shows a strong correlation with the efficiency of indexing technique and can provide a good reference on the necessity of building the index. The results of extensive experiments on two large real-world datasets demonstrate the effectiveness and efficiency of the proposed method. %With the support of \kw{MinLevel} strategy, the indexes are able to achieve good acceleration effect and space efficiency, especially segment tree, which also has the best scalability.
	\end{abstract}

%	\begin{abstract}
%		\justifying We study the problem of enabling data warehouse and OLAP (Online Analytical Processing) technology to analyze large scale temporal multidimensional networks. Data warehouse and OLAP are effective tools for decision support on traditional relational data and static multidimensional networks. However, many real-world multidimensional networks are often modeled as temporal multidimensional networks, where each edge in the network is associated with temporal information. Such temporal multidimensional networks often cannot be handled by traditional data warehouse and OLAP techniques. To fill this gap, we propose a novel data warehouse model, \kw{Temporal\enspace Graph\enspace Cube}. By extending the basic idea of static graph cube, \kw{Temporal\enspace Graph\enspace Cube} supports OLAP queries on temporal multidimensional networks. Specifically, by supporting OLAP queries in any time range, users can obtain summarized information of the networks in the time range of interest, which is impossible with traditional static graph cube. In conducting OLAP queries of \kw{Temporal\enspace Graph\enspace Cube}, we need to merge snapshots in a certain time range. We model this problem as a range query problem on snapshot arrays, and generalize three indexes in traditional range query problem: prefix array, sparse table and segment tree to accelerate the query. We also develop new updating techniques to maintain these generalized indexes when the temporal multidimensional network evolves over time. To further improve the query efficiency, we propose a new concept called \kw{similarity\enspace of\enspace snapshots} which shows a strong correlation with the efficiency of indexing techniques and can provide a good reference on the necessity of building the indexes. The results of extensive experiments on both real-world and synthetic datasets show the efficiency and effectiveness of the proposed solutions. %With the support of \kw{MinLevel} strategy, the indexes are able to achieve good acceleration effect and space efficiency, especially segment tree, which also has the best scalability.
%	\end{abstract}

	% Note that keywords are not normally used for peerreview papers.
%	\begin{IEEEkeywords}
%		Uncertain Graphs, Cohesive Subgraph Mining, Uncertain Core Decomposition.
%\end{IEEEkeywords}
}

@@

Data warehouses and OLAP (Online Analytical Processing) techniques are helpful tools for knowledge workers (executive, manager, analyst) to analyze and make decisions, because Data warehouses and OLAP together can efficiently provide summarized information in different resolutions by specifying different views (different combinations of data dimensions) of large-scale real-world data through OLAP operations such as roll-up, drill-down and slice-and-dice \cite{chaudhuri1997overview}. Data warehouse built on traditional relational database (RDB) data is called data cube \cite{gray1997data}. In 2011, Zhao et al.\ \cite{zhao2011graph} proposed \kw{Graph} \kw{Cube} which extends data warehouses and OLAP techniques to analyze static multidimensional networks. For each query, static graph cube returns a static network with summarized information in its structure and statistical values on vertices and edges.

However, many real-life networks, such as human proximity networks, scientific collaboration networks, and biological networks, can be modeled as temporal networks \cite{holme2015modern}, where each relationship or interaction has a timestamp. Also, vertices in temporal networks usually contain attributes of multiple dimensions. For example, vertices in human proximity networks have attributes such as name, gender, nationality, hobbies, and so on. Such an attributed network can be modeled as a temporal multidimensional network. There are many studies focusing on the management and analysis of temporal networks \cite{junghanns2017management}, but none of them have tried to extend OLAP techniques to temporal multidimensional networks, i.e., developing approaches to provide users in real time with summarized information on the temporal multidimensional networks in different resolutions and time ranges.
\begin{figure*}[t!]
	\begin{center}
	\begin{tabular}[t]{c}
		\centering
		\subfigure[{\scriptsize Vertex table}]{
			\includegraphics[width=0.27\textwidth]{fig/example/example_tables/tb_example_dataset_adjust.pdf}
		}
		\subfigure[{\scriptsize Network structure}]{
			\includegraphics[width=0.63\textwidth]{fig/example/example_dataset_trade.pdf}
		}
	\end{tabular}
	\end{center}
	\vspace*{-0.4cm}
	\caption{Running example: a temporal multidimensional network with its vertex table and network structure}
	\vspace*{-0.4cm}
	\label{fig:example_dataset}
\end{figure*}

\begin{example} \label{exam:temp_multi_network}
	Fig.~\ref{fig:example_dataset} shows a sample temporal transaction network, presenting the transactions between individuals of different countries of birth and professions in each day. Table in Fig.~\ref{fig:example_dataset} (a) is the vertex table of the temporal network, showing the information of the 8 individuals. Each individual has a primary key ID and 3 dimensions (3 discrete attributes): gender, country (country of birth), profession. Income is a numeric attribute of individuals. Fig.~\ref{fig:example_dataset} (b) shows the network structure of the temporal network. There are 20 temporal edges, illustrating the transactions between individuals in 5 days. Temporal edges in each day can be organized into a snapshot. The numeric attribute on each temporal edge is the amount of each transaction. The static multidimensional information of individuals in Fig.~\ref{fig:example_dataset} (a) and the temporal information of network structure in Fig.~\ref{fig:example_dataset} (b) form a temporal multidimensional network.
\end{example}
In static graph cube \cite{zhao2011graph}, users can specify different views to obtain summarized information of the static multidimensional network and combine the information under these views through OLAP operations such as roll-up, drill-down and slice-and-dice interactively for decision support and business intelligence. For example, a company tries to use static graph cube to analyze the interaction characteristics of people with different attributes (name, gender, profession, income level, etc.) in a large-scale social network in order to support their marketing strategy making. However, edges in real-world networks usually contain timestamps. For example, in social networks, interactions between people occur at specific times, such as a particular day. In this way, static graph cube does not work when users only need information for a specific time range. Considering the previous example, the company only needs data from the last few years, because data that is too old is less informative. Or, they want to study what the characteristics of people's interactions are during particular time ranges, such as holidays or COVID-19 periods. Or, they wish to compare several different time ranges to see how people's interaction patterns change over time. Static graph cube cannot meet the above requirements. As a result, we need to propose novel approaches to make data warehouses and OLAP capable of analyzing temporal multidimensional networks.

\comment{
\begin{figure}[t!]
	\begin{center}
		\begin{tabular}[t]{c}
			\centering
			\hspace*{+0.3cm}
			\subfigure[{\scriptsize Vertex table}]{
				\includegraphics[width=0.36\columnwidth]{fig/example/example_tables/tb_cuboid_gender_adjust.pdf}
			}
			\hspace*{-0.6cm}
			\subfigure[{\scriptsize Network structure}]{
				\includegraphics[width=0.58\columnwidth]{fig/example/exam_cuboid_gender_2_3.pdf}
			}
			\hspace*{-0.6cm}
		\end{tabular}
	\end{center}
	\vspace*{-0.4cm}
	\caption{The answer to ``What is the amount of transactions between individuals of different genders from day 2 to day 3?"}
	\vspace*{-0.4cm}
	\label{fig:exam_cuboid_gender}
\end{figure}
\begin{figure}[t!]
	\begin{center}
		\includegraphics[width=0.45\columnwidth]{fig/example/exam_cuboid_gender_semi_state.pdf}
	\end{center}
	\vspace*{-0.7cm}
	\caption{Intermediate result of generating network structure in Fig.~\ref{fig:exam_cuboid_gender} (b)}
	\vspace*{-0.4cm}
	\label{fig:exam_inter_result_gender}
\end{figure}

}

@@

%\begin{figure}[t!]
%	\begin{center}
%		\begin{tabular}[t]{c}\hspace*{-0.3cm}
%			\subfigure[{\scriptsize Structure of the answer of the first question in Example~\ref{exam:cuboid_query}}]{
%				\includegraphics[width=0.45\columnwidth]{fig/example/exam_cuboid_gender_2_3_adjust.pdf}
%			}
%			\subfigure[{\scriptsize The intermediate result of generating (a)}]{
%				\includegraphics[width=0.45\columnwidth]{fig/example/exam_cuboid_gender_semi_state.pdf}
%			}
%		\end{tabular}
%	\end{center}
%	\vspace*{-0.4cm}
%	\caption{}
%	\vspace*{-0.4cm}
%	\label{fig:exam_cuboid_gender}
%\end{figure}

\comment{
\begin{figure}[t!]
	\begin{center}
		\begin{tabular}[t]{c}\hspace*{-0.2cm}
			\subfigure[{\scriptsize Vertex table}]{
				\includegraphics[width=0.4\columnwidth]{fig/example/example_tables/tb_cuboid_gender_pro_adjust.pdf}
			}
			\subfigure[{\scriptsize Network structure from day 2 to day 3}]{
				\includegraphics[width=0.5\columnwidth]{fig/example/exam_cuboid_gender_pro_2_3.pdf}
			}
			%\vspace*{-0.3cm}\\\hspace*{-0.3cm}
			\\
			\subfigure[{\scriptsize Network structure from day 4 to day 5}]{
				\includegraphics[width=0.5\columnwidth]{fig/example/exam_cuboid_gender_pro_4_5.pdf}
			}
		\end{tabular}
	\end{center}
	\vspace*{-0.4cm}
	\caption{The answers to ``What is the amount of transactions between individuals of different genders and professions from day 2 to day 3 and from day 4 to day 5?"}
	\vspace*{-0.4cm}
	\label{fig:exam_cuboid_gender_profession}
\end{figure}
\begin{example} \label{exam:cuboid_query}
	Fig.~\ref{fig:exam_cuboid_gender} shows a static aggregate network obtained by summarizing the temporal multidimensional network in Fig.~\ref{fig:example_dataset} on dimension ``Gender'' and time range $ [2,3] $. Fig.~\ref{fig:exam_cuboid_gender} can be regarded as the answer of question ``What is the amount of transactions between individuals of different genders from day 2 to day 3?" Note that the vertex table in Fig.~\ref{fig:exam_cuboid_gender} (a) is directly aggregated from the vertex table in Fig.~\ref{fig:example_dataset} (a) on ``Gender" with COUNT($\ast$) as aggregation function on vertices. All vertices in Fig.~\ref{fig:example_dataset} (a) are included in the aggregation process even those vertices which have no adjacent edge in day 2 and day 3, like $ v_5 $. The reason is that vertices are time independent and vertices having no adjacent edge in specified time ranges are still available in those time ranges. Therefore, from day 2 to day 3, the COUNT($\ast$) value of male should be 4. Fig.~\ref{fig:exam_inter_result_gender} shows the intermediate result of the question, which is obtained by merging all snapshots in day 2 and day 3. All edges of snapshots in day 2 and day 3 are collected to the single snapshot in Fig.~\ref{fig:exam_inter_result_gender}, and if there are two edges in snapshots of day 2 and day 3 respectively share the same vertices, then their numeric attributes will be added together since SUM($\ast$) is the aggregation function upon edges here. Fig.~\ref{fig:exam_cuboid_gender} (b) can be directly aggregated from Fig.~\ref{fig:exam_inter_result_gender} on the basis of aggregated vertices in Fig.~\ref{fig:exam_cuboid_gender} (a), i.e., each edge in Fig.~\ref{fig:exam_cuboid_gender} (b) is the aggregated edge of edges in Fig.~\ref{fig:exam_inter_result_gender}, whose two vertices are aggregated to the two vertices of the corresponding aggregated edge. The numeric attributes of the aggregated edges in Fig.~\ref{fig:exam_cuboid_gender} (b) are still the sum of attributes of their corresponding sets of edges in Fig.~\ref{fig:exam_inter_result_gender}.

	Finally, we can reply the question above. For example, from day 2 to day 3, the amount of transactions between males is 6, while the amount of transactions between male and female is 107, etc. Fig.~\ref{fig:exam_cuboid_gender_profession} shows the answers of two similar questions ``What is the amount of transactions between individuals of different genders and professions from day 2 to day 3 and from day 4 to day 5?" The only difference between these two questions is the time range specified. Both static aggregate networks in Fig.~\ref{fig:exam_cuboid_gender_profession} share the same vertex table in Fig.~\ref{fig:exam_cuboid_gender_profession} (a) and in these two questions the aggregation functions upon vertices are AVERAGE($\ast$).
\end{example}

There are three new OLAP queries in Example~\ref{exam:cuboid_query} compared to queries in static graph cube, for example, the first OLAP query in Example~\ref{exam:cuboid_query} is ``What is the amount of transactions between individuals of different genders from day 2 to day 3", which requests a rough summarized result of the temporal multidimensional network in Fig.~\ref{fig:example_dataset} on ``Gender". Unlike the query in static graph cube, this new query includes a specific time range $ [2,3] $. Conducting the new query includes the same aggregation on vertex table as in static graph cube, but when aggregating the network structure, the basic idea is to first merge snapshots in the time range to build an intermediate result, as shown in Fig.~\ref{fig:exam_inter_result_gender}, then perform the same aggregation on Fig.~\ref{fig:exam_inter_result_gender} as in static graph cube. If the new query specifies the whole time range of the temporal multidimensional network, then it is equivalent to a query in static graph cube. Therefore, static graph cube is a special case of the new data warehouse model proposed in this paper. Conducting the second OLAP query ``What is the amount of transactions between individuals of different genders and professions from day 2 to day 3" can be regarded as performing a drill-down operation on the basis of the first query, since the second query requests more fine-grained information in the same time range than the first query. The third query ``What is the amount of transactions between individuals of different genders and professions from day 4 to day 5" specifies the same dimensions as the second query, but specifies a different time range $ [4,5] $. This allows users to analyze the pattern characteristics in different time ranges, for example, there is no transaction between male students from day 2 to day 3, but there are transactions totaling 100 between them from day 4 to day 5. Note that the aggregation functions upon vertices and edges can be combined arbitrarily as needed. For example, we can specify AVERAGE($\ast$) as the aggregation function upon both vertices and edges in Example~\ref{exam:cuboid_query}, so that we can analyze the transaction network from another perspective.

In \cite{zhao2011graph}, Zhao et al.\ proposed a new OLAP query for static multidimensional networks, called \kw{crossboid}, to analyze the crossed interaction pattern of two different views. It is easy to see that \kw{Crossboid} can also be extended to temporal multidimensional networks as illustrated in Example~\ref{exam:crossboid_query}.

>>>>

\documentclass[10pt,journal,compsoc]{IEEEtran}
%\usepackage{ntheorem}
%\usepackage{algorithmic}
%\usepackage{cite}
%\usepackage{times}
\usepackage{amsmath,amsfonts,amssymb, amsthm, xspace}
\usepackage[linesnumbered,ruled,vlined]{algorithm2e}
\usepackage{graphicx, subfigure,stfloats}
\usepackage{epstopdf}
\usepackage{textcomp}
\usepackage{ragged2e, xcolor}
\usepackage{balance}
\usepackage{verbatim}
\usepackage{enumitem}
\usepackage{makecell}
\usepackage{url}
\usepackage{booktabs}
%
% If IEEEtran.cls has not been installed into the LaTeX system files,
% manually specify the path to it like:
% \documentclass[10pt,journal,compsoc]{../sty/IEEEtran}

@@

\renewcommand{\baselinestretch}{0.99}

@@

	\author{Qiangqiang~Dai, Rong-Hua~Li, Guoren~Wang, Rui~Mao, Zhiwei~Zhang, Ye~Yuan
		\IEEEcompsocitemizethanks{
			\IEEEcompsocthanksitem Q.~Dai, R.-H.~Li, G.~Wang, Z.~Zhang and Y.~Yuan are with the Beijing Institute of Technology, Beijing 100081, China. E-mail: qiangd66@gmail.com, lironghuascut@gmail.com, wanggrbit@126.com, cszwzhang@outlook.com, yuanye@mail.neu.edu.cn.
			\IEEEcompsocthanksitem R.~Mao is with the Shenzhen University, Shenzhen 518060, China. E-mail: mao@szu.edu.cn.
		}
	}

@@

% The publisher's ID mark at the bottom of the page is less important with
% Computer Society journal papers as those publications place the marks
% outside of the main text columns and, therefore, unlike regular IEEE
% journals, the available text space is not reduced by their presence.
% If you want to put a publisher's ID mark on the page you can do it like
% this:
%\IEEEpubid{0000--0000/00\$00.00~\copyright~2015 IEEE}
% or like this to get the Computer Society new two part style.
%\IEEEpubid{\makebox[\columnwidth]{\hfill 0000--0000/00/\$00.00~\copyright~2015 IEEE}%
	%\hspace{\columnsep}\makebox[\columnwidth]{Published by the IEEE Computer Society\hfill}}
% Remember, if you use this you must call \IEEEpubidadjcol in the second
% column for its text to clear the IEEEpubid mark (Computer Society jorunal
% papers don't need this extra clearance.)

@@

% for Computer Society papers, we must declare the abstract and index terms
% PRIOR to the title within the \IEEEtitleabstractindextext IEEEtran
% command as these need to go into the title area created by \maketitle.
% As a general rule, do not put math, special symbols or citations
% in the abstract or keywords.
\IEEEtitleabstractindextext{%
	\begin{abstract}
		\justifying We study the problem of enabling data warehouse and OLAP technology to analyze large scale temporal multidimensional networks. Data warehouse and OLAP (Online Analytical Processing) are effective tools for decision support on traditional relational data and static multidimensional network. However, real-world networks can usually be modeled as temporal multidimensional networks and they can not be handled by data warehouse and OLAP technology yet. In this paper, we propose a new data warehouse model, \kw{Temporal\enspace Graph\enspace Cube}. By extending the basic idea of static graph cube, \kw{Temporal\enspace Graph\enspace Cube} supports OLAP queries on temporal multidimensional networks. Specifically, by supporting OLAP queries in any time range, users can obtain summative information of the networks in the time range of interest, which is impossible with static graph cube. In conducting OLAP queries of \kw{Temporal\enspace Graph\enspace Cube}, we should merge snapshots in a certain time range. We abstract this problem as range query problem on snapshot array, and we extend three typical indexes in commong range query problem: prefix array, sparse table and segment tree to accelerate the query. We enable these indexes to be updated in real time, because the size of real-world temporal networks grow over time. We find that the above indexes are not always work, and sometimes the baseline method is the best method. Based on this, we define \kw{similarity\enspace of\enspace snapshots}, which is proved to have a strong correlation with the efficiency of indexes and can provide a reference on the necessity to build indexes. In implementing \kw{Temporal\enspace Graph\enspace Cube}, we find that the traditional algorithms used to implement traditional data cube and static graph cube are no longer applicable to \kw{Temporal\enspace Graph\enspace Cube}. Instead, we adopt another simple method \kw{MinLevel}, which is introduced in implementation of static graph cube. The results of extensive experiments on real-world and man-made datasets show the effectiveness of \kw{Temporal\enspace Graph\enspace Cube} and corroborate the correlation between \kw{similarity\enspace of\enspace snapshots} and the efficiency of indexes. With the support of \kw{MinLevel} strategy, the indexes are able to achieve good acceleration effect and space efficiency, especially segment tree, which also has the best scalability.
	\end{abstract}

	% Note that keywords are not normally used for peerreview papers.
	\begin{IEEEkeywords}
		Uncertain Graphs, Cohesive Subgraph Mining, Uncertain Core Decomposition.
\end{IEEEkeywords}}

@@

Data warehouses and OLAP (Online Analytical Processing) technology have proven to be helpful tools for knowledge workers (executive, manager, analyst) to analyze and make decisions, because Data warehouses and OLAP together can efficiently provide summarized information in different resolutions by specifying different views (different combinations of data dimensions) of large-scale real-world data through OLAP operations such as roll-up, drill-down and slice-and-dice \cite{chaudhuri1997overview}. Data warehouse built on traditional RDB data can be called data cube \cite{gray1997data}. Zhao et al. \cite{zhao2011graph} proposed \kw{Graph} \kw{Cube} in 2011 which extends data warehouses and OLAP techniques so that they can be used to analyze static multidimensional networks. For each query, static graph cube returns a static network, and the summative information is in it's structure and statistical values on vertices and edges, unlike traditional data cube which only returns the statistical values obtained by group-by operations. However, many real-life networks, such as human proximity networks, scientific collaboration networks, and biological networks can be modeled as temporal networks \cite{holme2015modern}, where each relationship or interaction has a corresponding timestamp. Also, vertices in temporal networks usually contain attributes of multiple dimensions, for example, vertices in human proximity networks have attributes such as name, gender, nationality, hobbies, etc. Therefore, we can further construct the above networks as temporal multidimensional networks. There has been many works focusing on the management and analysis of temporal networks \cite{junghanns2017management}, but none of them have tried to extend OLAP techniques to temporal multidimensional networks, i.e., developing ways to provide users in real time with summative information about temporal multidimensional networks in different resolutions and time ranges.
\begin{figure*}[t!]
	\begin{center}
	\begin{tabular}[t]{c}
		\centering
		\subfigure[{\scriptsize Vertex table}]{
			\includegraphics[width=0.27\textwidth]{fig/example/example_tables/tb_example_dataset_adjust.pdf}
		}
		\subfigure[{\scriptsize Network structure}]{
			\includegraphics[width=0.63\textwidth]{fig/example/example_dataset_trade.pdf}
		}
	\end{tabular}
	\end{center}
	\vspace*{-0.4cm}
	\caption{A sample temporal multidimensional network with its vertex table and network structure}
	\vspace*{-0.4cm}
	\label{fig:example_dataset}
\end{figure*}

\begin{example} \label{exam:temp_multi_network}
	Figure~\ref{fig:example_dataset} shows a sample temporal transaction network, presenting the transactions between individuals of different countries of birth and professions in each day. Table in Figure~\ref{fig:example_dataset} (a) is the vertex table of the temporal network, showing the information of the 8 individuals. Each individual has a primary key ID and 3 dimensions (3 discrete attributes): gender, country (country of birth), profession. Income is a numeric attribute of individuals. Figure~\ref{fig:example_dataset} (b) shows the network structrue of the temporal network. There are 20 temporal edges, illustrating the transactions between individuals in 5 days. Temporal edges in each day can be organized into a snapshot. The numeric attribute on each temporal edge is the amount of each transaction. The static multidimensional information of individuals in Figure~\ref{fig:example_dataset} (a) and the temporal information of network structure in Figure~\ref{fig:example_dataset} (b) form a temporal multidimensional network.
\end{example}
In static graph cube, users can specify different views to obtain summative information about the static multidimensional network and combine the information under these views through OLAP operations such as roll-up, drill-down and slice-and-dice interactively for decision support and business intelligence. For example, a company tries to use static graph cube to analyze the interaction characteristics of people with different attributes (name, gender, profession, income level, etc.) in a large-scale social network in order to support their marketing strategy making. However, edges in real-world networks usually contain timestamps. For example, in social networks, interactions between people occur at specific times, such as a particular day. In this way, static graph cube does not work when users only need information for a specific time range. Taking the previous example, the company only needs data from the last few years, because data that is too old is less informative. Or, they want to study what the characteristics of people's interactions are during particular time ranges, such as holidays or COVID-19 periods. Or, they want to compare several different time ranges to see how people's interaction patterns change over time. Static graph cube can not meet the above needs. As a result, we need to extend static graph cube to make data warehouses and OLAP capable of analyzing temporal multidimensional networks.

\begin{figure}[t!]
	\begin{center}
		\begin{tabular}[t]{c}
			\centering
			\hspace*{+0.3cm}
			\subfigure[{\scriptsize Vertex table}]{
				\includegraphics[width=0.36\columnwidth]{fig/example/example_tables/tb_cuboid_gender_adjust.pdf}
			}
			\hspace*{-0.6cm}
			\subfigure[{\scriptsize Network structure}]{
				\includegraphics[width=0.58\columnwidth]{fig/example/exam_cuboid_gender_2_3.pdf}
			}
			\hspace*{-0.6cm}
		\end{tabular}
	\end{center}
	\vspace*{-0.4cm}
	\caption{The answer to ``What is the amount of transactions between individuals of different genders from day 2 to day 3?"}
	\vspace*{-0.4cm}
	\label{fig:exam_cuboid_gender}
\end{figure}
\begin{figure}[t!]
	\begin{center}
		\includegraphics[width=0.45\columnwidth]{fig/example/exam_cuboid_gender_semi_state.pdf}
	\end{center}
	\vspace*{-0.7cm}
	\caption{Intermediate result of generating network structure in Figure~\ref{fig:exam_cuboid_gender} (b)}
	\vspace*{-0.4cm}
	\label{fig:exam_inter_result_gender}
\end{figure}

@@

%\begin{figure}[t!]
%	\begin{center}
%		\begin{tabular}[t]{c}\hspace*{-0.3cm}
%			\subfigure[{\scriptsize Structrue of the answer of the first question in Example~\ref{exam:cuboid_query}}]{
%				\includegraphics[width=0.45\columnwidth]{fig/example/exam_cuboid_gender_2_3_adjust.pdf}
%			}
%			\subfigure[{\scriptsize The intermediate result of generating (a)}]{
%				\includegraphics[width=0.45\columnwidth]{fig/example/exam_cuboid_gender_semi_state.pdf}
%			}
%		\end{tabular}
%	\end{center}
%	\vspace*{-0.4cm}
%	\caption{}
%	\vspace*{-0.4cm}
%	\label{fig:exam_cuboid_gender}
%\end{figure}

\begin{figure}[t!]
	\begin{center}
		\begin{tabular}[t]{c}\hspace*{-0.2cm}
			\subfigure[{\scriptsize Vertex table}]{
				\includegraphics[width=0.4\columnwidth]{fig/example/example_tables/tb_cuboid_gender_pro_adjust.pdf}
			}
			\subfigure[{\scriptsize Network structure from day 2 to day 3}]{
				\includegraphics[width=0.5\columnwidth]{fig/example/exam_cuboid_gender_pro_2_3.pdf}
			}
			%\vspace*{-0.3cm}\\\hspace*{-0.3cm}
			\\
			\subfigure[{\scriptsize Network structure from day 4 to day 5}]{
				\includegraphics[width=0.5\columnwidth]{fig/example/exam_cuboid_gender_pro_4_5.pdf}
			}
		\end{tabular}
	\end{center}
	\vspace*{-0.4cm}
	\caption{The answers to ``What is the amount of transactions between individuals of different genders and professions from day 2 to day 3 and from day 4 to day 5?"}
	\vspace*{-0.4cm}
	\label{fig:exam_cuboid_gender_profession}
\end{figure}
\begin{example} \label{exam:cuboid_query}
	Figure~\ref{fig:exam_cuboid_gender} shows a static aggregate network obtained by summarizing the temporal multidimensional network in Figure~\ref{fig:example_dataset} on dimension ``Gender'' and time range $ [2,3] $. Figure~\ref{fig:exam_cuboid_gender} can be regarded as the answer of question ``What is the amount of transactions between individuals of different genders from day 2 to day 3?" Note that the vertex table in Figure~\ref{fig:exam_cuboid_gender} (a) is directly aggregated from the vertex table in Figure~\ref{fig:example_dataset} (a) on ``Gender" with COUNT($\ast$) as aggregation function on vertices. All vertices in Figure~\ref{fig:example_dataset} (a) are included in the aggregation process even those vertices which have no adjacent edge in day 2 and day 3, like $ v_5 $. The reason is that vertices are time independent and vertices having no adjacent edge in specified time ranges are still available in those time ranges. So from day 2 to day 3, the COUNT($\ast$) value of male should be 4. Figure~\ref{fig:exam_inter_result_gender} shows the intermediate result of the question, which is obtained by merging all snapshots in day 2 and day 3. All edges of snapshots in day 2 and day 3 are collected to the single snapshot in Figure~\ref{fig:exam_inter_result_gender}, and if there are two edges in snpshots of day 2 and day 3 respectively share the same vertices, then their numeric attributes will be added together since SUM($\ast$) is the aggregation function upon edges here. Figure~\ref{fig:exam_cuboid_gender} (b) can be directly aggregated from Figure~\ref{fig:exam_inter_result_gender} on the basis of aggregated vertices in Figure~\ref{fig:exam_cuboid_gender} (a), i.e., each edge in Figure~\ref{fig:exam_cuboid_gender} (b) is the aggregated edge of edges in Figure~\ref{fig:exam_inter_result_gender}, whose two vertices are aggregated to the two vertices of the corresponding aggregated edge. The numeric attributes of the aggregated edges in Figure~\ref{fig:exam_cuboid_gender} (b) are still the sum of attributes of their corresponding sets of edges in Figure~\ref{fig:exam_inter_result_gender}.

	Finally we can reply the question above, for example, from day 2 to day 3, the amount of transactions between males is 6, while the amount of transactions between male and female is 107, etc. Figure~\ref{fig:exam_cuboid_gender_profession} are the answers of two similar questions ``What is the amount of transactions between individuals of different genders and professions from day 2 to day 3 and from day 4 to day 5?" The only difference between these two questions is the time range specified. Both static aggregate networks in Figure~\ref{fig:exam_cuboid_gender_profession} share the same vertex table in Figure~\ref{fig:exam_cuboid_gender_profession} (a) and in these two questions the aggregation functions upon vertices are AVERAGE($\ast$).
\end{example}
There are three new OLAP queries in Example~\ref{exam:cuboid_query} compared to queries in static graph cube, for example, the first OLAP query in Eample~\ref{exam:cuboid_query} is ``What is the amount of transactions between individuals of different genders from day 2 to day 3", which requests a rough summative result of the temporal multidimensional network in Figure~\ref{fig:example_dataset} on ``Gender". Unlike the query in static graph cube, this new query includes a specific time range $ [2,3] $. Conducting the new query includes the same aggregation on vertex table as in static graph cube, but when aggregating the network structure, the basic idea is to first merge snapshots in the time range to build an intermediate result, as in Figure~\ref{fig:exam_inter_result_gender}, then perform the same aggregation on Figure~\ref{fig:exam_inter_result_gender} as in static graph cube. If the new query specifies the whole time range of the temporal multidimensional network, then it's equivalent to a query in static graph cube. Therefore, static graph cube is a special version of the new data warehouse model to be proposed in this paper. Conducting the second OLAP query ``What is the amount of transactions between individuals of different genders and professions from day 2 to day 3" can be regarded as performing a drill-down operation on the basis of the first query, since the second query requests more fine-grained information in the same time range than the first query. The third query ``What is the amount of transactions between individuals of different genders and professions from day 4 to day 5" specifies the same dimensions as the second query, but specifies a different time range $ [4,5] $. This allows users to analyze the pattern characteristics in different time ranges, for example, there is no transaction between male students from day 2 to day 3, but there are transactions totaling 100 between them from day 4 to day 5. Note that the aggregation functions upon vertices and edges can be combined arbitrarily as needed. For example, we can specify AVERAGE($\ast$) as the aggregation function upon both vertices and edges in Example~\ref{exam:cuboid_query}, so that we can analyze the transaction network from another perspective.

In \cite{zhao2011graph}, Zhao et al. proposed a new OLAP query especially for static multidimensional networks \kw{crossboid}, to analyze the crossed interaction pattern of two different views. \kw{Crossboid} can also be easily extended to the temporal multidimensional network scenario as in Example~\ref{exam:crossboid_query}.

\begin{table}[t!]%\vspace*{-0.3cm}
	%\scriptsize
	\small
	\centering
	\caption{Vertex table obtained by group-by on ``Profession" in Figure~\ref{fig:example_dataset} (a)} \vspace*{-0.2cm} \label{tab:groupby_profession}
	\begin{tabular}{c c}
		\toprule
		Profession & COUNT($\ast$) \cr %\hline %\hline
		\midrule
		teacher& 1 \cr %\hline
		shopowner & 2  \cr %\hline
		student & 2  \cr %\hline
		lawyer & 2  \cr %\hline
		doctor & 1  \cr %\hline
		\bottomrule
	\end{tabular}
	\vspace*{-0.4cm}
\end{table}
